Experiences designing and building a PRAGMA Cloud Scheduler

### 2015-08-14 paper.tex
Added last and first sections

\maketitle
\begin{abstract}
PRAGMA is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas of biodiversity distribution and lake ecology.  Over the past four years, the technology focus for PRAGMA partners has shifted to cloud and software defined networking as enabling technologies.  During PRAGMA 27 meeting, the Resources Working Group discussed rebooting the persistent PRAGMA Testbed as a way for sites to contribute and use shared resources, leveraging technologies such as PRAGMA Boot, Personal Cloud Controller (PCC), and virtual network overlays (i.e. IPOP and ViNe).  To make PRAGMA resource sharing easier, a lightweight scheduler was proposed and discussed as a way to enable access to the resources and to manage resources reservations.  This short paper discusses the design process and initial prototype of a simple cloud scheduler for PRAGMA.

@@

We next looked more broadly for a simpler approach to scheduling and found a number of Web-based calendar reservation systems that seemed promising.  These systems were designed to manage meeting room reservations, hotel room reservations, or equipment rentals.   Their interfaces would be also be familiar to users in their everyday lives and thus well suited to the different PRAGMA users who have varying levels of distributed systems and networking expertise.  Also, several of these tools were significantly less complex than the batch and testbed scheduling tools above.  Later, we found a Web-based calendar reservation system called GridPrems~\cite{gridprems} for Grid'5000 indicating a simpler interface was sometimes needed even for their users.

%Some of the calendar reservation systems we looked at like Reservations~\cite{drupalreservations}, Merci~\cite{merci}, and Booking\_calendar~\cite{wordpressbooking} were plugins for Drupal~\cite{drupal} or Wordpress~\cite{wordpresss}.  However, the majority of tools were PHP based with a relational database backend like the Meeting Room Booking System~\cite{mrbs}, Booked~\cite{booked}, and Classroombookings~\cite{classroombookings}.

Since we were primarily interested in a scheduler with a like GridPrems, we decided to look at other open source calendar reservation systems that were available.  We looked at several systems used primarily for tracking room and/or equipment reservations and evaluated the following criteria:

\textbf{Open Source and customizable:}  Open source software is generally preferable because it provides it is free to use and can be customized as needed.  Since we did not know of any existing Web interfaces for managing virtual clusters, we limited our search to customizable Open Source tools and evaluated how easy it would be to manage resources and users as well as add new fields and features.

\textbf{Nice GUI interface:}  We evaluated the GUI interfaces of several calendar reservation tools and looked for those that had intuitive menus and navigation and  a clean and uncluttered look.  Some of the GUI interfaces we looked at just used basic HTML tags and looked very clunky so we limited our search to tools that leveraged CSS and Javascript and looked more professional.

\textbf{Installation and maintenance:}  We downloaded a number of tools and reviewed their installation instructions to see how easy they were to setup a demo instance.  Most of the tools we looked at were developed in PHP and had a relational database backend that was could be served from a basic  Apache web server\~cite{apache} setup.   A few of the tools were plugins for Drupal.  However, we eliminated a number of tools because they had either too little documentation or complex installation instructions with several manual steps.

Discuss selection of Booked.

Open source
Easy to setup
Nice GUI interface
Report features
REST API
Customizable-ish
LDAP and Active Directory support.
Fine tuned roles and permissions.
User and group quotas.

Can only handle one reservation per resource at a time
PHP changes can be painful (heavy OO makes it hard to find right files)
Doc is sparse

http://mrbs.sourceforge.net
	PHP, Postgres/MySQL
	Demo was a bit slow and a bit clunky looking
Booked - http://www.twinkletoessoftware.com/products
	Nice interface, lot of customization features
	PHP, MySQL
	RESTFUL API
http://classroombookings.com/
	Nice interface
	PHP, MySQL
http://www.unitime.org/unitime\_intro.php
	Geeky looking, might have some good algorithms
	Interface not as smooth
	http://journal.code4lib.org/articles/2941
	Clunky looking
http://www.drupalrooms.com
	More for hotel web sites ? e.g., prices, etc.
https://groups.drupal.org/node/137544
	Merci - https://www.drupal.org/project/merci
		For equipment, etc.
	https://www.drupal.org/project/reservations
		Also seems pretty flexible

 tried a few tools  here is the summary:
1. OTA hotel management   http://rocks-86.sdsc.edu/hotelmis/setup/initialsetup.php
    appears to be just for the hotel, no easy login, no easy change. Not useful
2. Mybooking - in Malay and after changing the language and trying enenglishigsh i did not get far. -- hotel
3. Booking\_calendar - strictly room booking, very small in features, too much to rewrite. -- wordpress plugin
4 Optaplanner appears to be a java app that runs outside of web. I tried examples and i am
   not sure of its usability. The examples are running producing sometimes some output sometimes noe
   and they are just java apps on some data. The examples  are not impressive even in terms of giving
   an idea how things are run.

A few more links are on http://opensource.com/business/15/1/top-project-management-tools-2015
Started looking at  one of them, Tuleap, may be too big for what we need.

\section{Cloud Scheduler Pilot}
\label{Sec:Pilot}

@@

@@

cloud scheduler would leverage 
To create a lightweight scheduler that enabled users to start and manage virtual clusters on leveraged the pragma\_boot, PCC, and software-defined networking tools that had been developed as part of PRAGMA 
As 
Our 
To schedule virtual clusters on the PRAGMA cloud testbed, 
Discuss creation of Cloud Scheduler Pilot.

@@

Some interesting conclusion.

>>>>

\maketitle
\begin{abstract}
PRAGMA is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas of biodiversity distribution and lake ecology.  Over the past four years, the technology focus for PRAGMA's testbed has shifted to cloud and software defined networking as enabling technologies.  This short paper describes the design of a simple cloud scheduler built on top of a Web reservation system that will enable users to easily run and manage virtual clusters for their science.   It is also equally important that there be a low barrier to entry for PRAGMA institutions to become part of the PRAGMA cloud testbed and install the needed software.  Therefore, the cloud scheduler  has a simple client-server architecture and will be built upon existing technologies such as pragma\_boot, Personal Cloud Controller,  virtual network overlays (e.g.,. IPOP), and  and also a Web reservation system.  We discuss our experiences implementing our pilot cloud scheduler and then describe future work.

@@

We next looked more broadly for a simpler approach to scheduling and found a number of open source Web-based calendar reservation systems that seemed promising.  These systems were designed to manage meeting room reservations, hotel room reservations, or equipment rentals.   Their interfaces would be also be familiar to users in their everyday lives and thus well suited to the different PRAGMA users who have varying levels of distributed systems and networking expertise.  Also, several of these tools were significantly less complex than the batch and testbed scheduling tools above.  Later, we found a Web-based calendar reservation system called GridPrems~\cite{gridprems} for Grid'5000 indicating a simpler interface was sometimes needed even for their users.

Some of the calendar reservation systems we looked at like Reservations~\cite{drupalreservations}, Merci~\cite{merci}, and Booking\_calendar~\cite{wordpressbooking} were plugins for Drupal~\cite{drupal} or Wordpress~\cite{wordpress}.  However, the majority of tools were PHP based with a relational database backend like the Meeting Room Booking System~\cite{mrbs}, Booked~\cite{booked}, and Classroombookings~\cite{classroombookings}.  For each system, we evaluated 1) how easy it would be to manage resources, reservations, and users as well as add new fields and features, 2) the GUI interface with respect to intuitive menus/navigation as well as a clean, modern, and uncluttered look, and 3)  how easy it was to install and setup a demo instance. One deficiency that all systems had was that they only allowed one reservation per resource.

Out of the calendar reservation systems we evaluated, Booked~\cite{booked} from Twinkle Toes software seemed the best suited.  It was easy to setup, had  a nice looking GUI interface, and was customizable.  It also had some nice additional features like usage reporting, a REST API interface, LDAP and Active Directory support, fine grained roles and permissions, and user/group quotas.  After working with Booked during our pilot, we did find some weaknesses.  First, the documentation was a bit sparse in areas and any needed PHP changes were time consuming due to its heavily object-oriented structure (i.e., we had to go thru many files to find the right place to change the code).

\section{Scheduler Architecture}
\label{Sec:Arch}

@@

\section{Scheduler Pilot}
\label{Sec:Pilot}

@@

@@

This paper described the the design and pilot implementation of a  scheduler for the PRAGMA cloud testbed.  Since the scheduler does not need to scale beyond tens of users, we prioritized ease of use and low installation/maintenance overhead and opted for a simple client-server design.  The server has an intuitive Web GUI frontend based on the Booked Web reservation system software.

The pilot implementation of the cloud scheduler made a lot of temporary assumptions and so our future work addresses them.  We will first enhance the Booked software to handle multiple reservations and then work on integrating IPOP and PRAGMA-ENT into the cloud scheduler as well.  Next, we will rework the PCC software and integrate HTCondor so it's used in personal mode allowing us to keep a light footprint on each of the PRAGMA sites.  Finally, we will investigate using the CloudFront option in pragma\_boot to manage application virtual cluster images and staging to each of the sites.     Documentation and packaging will also be done throughout so that new users and sites can get quickly onto the PRAGMA cloud testbed and use it as an experimental service to test new science and infrastructure tools..
